---
title: "Prep firm risk data"
author: "Gopal Penny"
date: "4/15/2021"
output: 
  html_document:
    toc: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Data background

The firm risk data is located [here](https://sites.google.com/view/firmrisk/home?authuser=0)

The dataset was developed by a group of researchers including:

* Tarek A. Hassan, Associate Professor of Economics, Boston University
* Stephan Hollander, Professor of Financial Accounting, Tilburg University
* Laurence van Lent, Professor of Accounting and Economics, Frankfurt School of Finance and Management
* Ahmed Tahoun, Associate Professor of Accounting, London Business School
* Markus Schwedeler, PhD student in Economics, Boston University

## Data prep

The original dataset had some funny quirks in the csv file, `firmquarter_2020q4.csv`. I re-named this as a text file (`firmquarter_2020q4.txt`), imported the data line-by-line, and then took only the first 22 columns. This resolved a couple of issues where `read.table` was not detecting line endings and some rows which seemed to have 35 columns instead of 36. 

```{r read-and-prep, eval = FALSE}
library(tidyverse)

# Read all lines of the file
con <- file("~/Downloads/firmquarter_2020q4.txt")
all_lines <- readLines(con)
close(con)

# The maximum number of characters in a line should be around 500. If it's greater, then line endings are not being parsed correctly
max(nchar(all_lines))

# Split each line into columns and save as individual data.frame
for (i in 1:length(all_lines)) {
  if (i == 1) {
    line_fields <- list(NULL)
  }
  line <- all_lines[i]
  fields <- strsplit(line, "\\t")
  line_fields[[i]] <- as.data.frame(t(fields[[1]]))
}

# Select the first 22 columns of each line
line_fields_select <- lapply(line_fields, function(x) x[1, 1:22])

# Combine all lines into a single data.frame
fq_prep <- do.call(rbind, line_fields_select[2:length(line_fields_select)]) #%>%

# Format the data.frame with column headers and convert numeric and date variables from character
fq <- fq_prep %>% as_tibble() %>%
  mutate(across(c(3:16), as.numeric),
         across(22, function(x) as.Date(x, "%d-%b-%Y"))) %>%
  set_names(line_fields_select[[1]])

# Save the data as an RDS file
saveRDS(fq, "data/format/firmquarter_2020q4.rds")


# Save the data as csv file
write_csv(fq, "data/format/firmquarter_2020q4.csv")
```

Ensure csv save is identical to original file:

```{r echo = FALSE}
fq_rds <- readRDS("data/format/firmquarter_2020q4.rds")

fq_csv <- read_csv("data/format/firmquarter_2020q4.csv") %>%
  mutate(ticker = replace(ticker, is.na(ticker),""))
identical(dim(fq_rds), dim(fq_csv)) # they are the same :)
```

```{r}
fq_num_rows <- nrow(fq)
fq_num_companies <- length(unique(fq$company_name))
```

The output data file `firmquarter_2020q4.rds` contains `r format(fq_num_rows,big.mark = ",")` rows of data corresponding to `r format(fq_num_companies,big.mark = ",")` companies.


## Plot data for a sample of firms

Sample 20 firms and plot the main data for risk and sentiment scores.

```{r fig.width = 10, fig.height = 13}
set.seed(100)
company_names_all <- unique(fq$company_name)
companies <- sample(company_names_all, 20)
fq_sample <- fq %>% filter(company_name %in% companies)

# Generate random colors
color <- grDevices::colors()[grep('gr(a|e)y', grDevices::colors(), invert = T)]
n <- 50
set.seed(50)
rand_colors <- sample(color, n)

# Prepare dates and 
fq_sample %>% pivot_longer(PRisk:PRiskT_technology) %>%
  # filter(name %in% c("Risk","Sentiment")) %>%
  ggplot(aes(date_earningscall, value, color = company_name)) + 
  geom_point(aes(shape = company_name)) + geom_line(alpha = 0.6) +
  scale_color_manual(values = rand_colors) + 
  scale_shape_manual(values = rep(1:6,10)) +
  facet_wrap(~name, scales = "free_y", ncol = 2) +
  ggp::t_manu()
```

